---
title: "Lab 2 Part 2 Solutions"
format: 
  html:
    toc: true
    toc-expand: 2
editor: visual
---

# Visualizing California COVID-19 hospital data

Today we are also going to work with California COVID hospital data available at https://data.ca.gov/dataset/covid-19-hospital-data.

## Question 1

**1a.** Begin by reading in the data by running the chunk below.

```{r}
library(tidyverse)

ca_covid_hosp_data <- read_csv("data/covid19hospitalbycounty.csv")
```

**1b.** Use `glimpse()` to get a preview of the data and view the columns.

**1b. Solution**

```{r}
glimpse(ca_covid_hosp_data)
```

**1c.** How many observations and variables are in this data set?
Which variables are categorical and which are numeric?

**1c. Solution**

There are 67,000 observations and 9 variables.
County and date are categorical while the rest are numeric.

## Question 2

**2a.** Run the following code chunk to make a bar plot of the number of hospitalized COVID confirmed patients by date and colored by county.

```{r}
ggplot(
  data = ca_covid_hosp_data,
  aes(
    x = todays_date,
    y = hospitalized_covid_confirmed_patients,
    fill = county
  )
) +   
  geom_bar(position = "stack", stat = "identity")
```

**2b.**

The above plot tells us something about CA trends, but prevents us from comparing trends among counties, in addition to being absolutely hideous.
Let's focus on just five counties.

```{r}
ca_five_county_covid_hosp_data <- ca_covid_hosp_data %>%   
  filter(county %in% c("Los Angeles", "Orange", "Sacramento", "Santa Clara", "San Francisco"))  
```

Redo the plot from question 2a with this subsetted data.

**2b. Solution**

```{r}
ggplot(   
  ca_five_county_covid_hosp_data,   
  aes(     
    fill = county,      
    x = todays_date,      
    y = hospitalized_covid_confirmed_patients   
  ) 
) +   
  geom_bar(position = "stack", stat = "identity")
```

## Question 3

The 5 county graph from question 2b is more readable than the graph in 2a, but still has tons of problems if one really wants to compare COVID-19 hospitalization trends across CA counties.
Create your own visualization of the 5 county data, remembering best practices that we talked about in the lecture.
There is more than one way of doing this, so don't be inhibited by trying to think of "the right" solution.

**3. Solution**

As we discussed, there is no one solution to this problem.
Visit [COVID-19 Trends by UC Irvine](https://www.stat.uci.edu/covid19/index.html) to view how our research team addressed this challenge.
