---
title: "linear mixed effects model"
format: html
editor: visual
---

## libraries

```{r}
library(tidyverse)
library(lme4)
```

## data {#data}

```{r}
chart_dat <- read_csv("chart_wide_condense.csv")

```

### Data Organization

Key variables in `chart_wide_condense.csv` which we will examine to address the research questions above are:

-   `schoolid` = includes district type, district number, and school number
-   `schoolName` = name of school
-   `urban` = is the school in an urban (1) or rural (0) location?
-   `charter` = is the school a charter school (1) or a non-charter public school (0)?
-   `schPctnonw` = proportion of non-white students in a school (based on 2010 figures)
-   `schPctsped` = proportion of special education students in a school (based on 2010 figures)
-   `schPctfree` = proportion of students who receive free or reduced lunches in a school (based on 2010 figures). This serves as a measure of poverty among school families.
-   `MathAvgScore.0` = average MCA-II math score for all sixth grade students in a school in 2008
-   `MathAvgScore.1` = average MCA-II math score for all sixth grade students in a school in 2009
-   `MathAvgScore.2` = average MCA-II math score for all sixth grade students in a school in 2010

This data is stored in WIDE format, with one row per school, as illustrated in Table \@ref(tab:table1chp9).

```{r, include=FALSE, warning=FALSE}
#Getting started-2
# Create data frame in LONG form (one obs per school-year)
#   chart.long is 1854x10 with 121 NAs for MathAvgScore
select <- dplyr::select
chart_long <- chart_dat %>%
  gather(key = "key", value = "MathAvgScore",
         MathAvgScore.0:MathAvgScore.2) %>%
  separate(key, into = c("name", "year08"), sep = "\\.") %>%
  select(-c("name")) %>%
  arrange(schoolid, year08) %>%
  mutate(year08 = as.numeric(year08))

```

## linear model

```{r}
model0 <- lm(MathAvgScore ~ year08, data = chart_long)
```

## fiting a simple model of time with a random intercept for school

```{r}
model1 <- lmer(MathAvgScore~ year08 + (1|schoolid), 
  REML=T, data= chart_long)
```

# plot the fitted values

## fitting a model of time with a random intercept and random slope for school

```{r}
model2 <- lmer(MathAvgScore~ year08 + (1 + year08|schoolid), 
  REML=T, data= chart_long)

```

# plot the fitted values

## adding in charter school

```{r}
model3 <- lmer(MathAvgScore~ charter + year08 + 
  charter:year08 + (year08|schoolid), 
  REML=T, data=chart_long)

```
